The Implementation of China’s Overseas NGO Law and the Operating Space for International Civil Society

### 2023-11-28 manuscript/manuscript.qmd
More cleanup + added TODOs for each remaining section

---
title: "The Implementation of China’s Overseas NGO Law and the Operating Space for International Civil Society"
short-title: "China's ONGO Law and International Civil Society"
date: "2023-11-27"
author:
- name: Meng Ye
  affiliation: Georgia State University
  email: mye2@student.gsu.edu
  orcid: 0000-0002-6552-8338
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
abstract: |
  China’s 2017 Overseas NGO (ONGO) Law is part of a larger global trend of legal restrictions on international nongovernmental organizations (INGOs). However, the effects of this global crackdown on INGOs have been difficult to measure since there is often divergence between the formal de jure regulations and the de facto implementation of these laws. Additionally, the severity and consistency of the enforcement of these laws often depends on the type of services provided by these organizations. In this paper, we explore the effect of China’s ONGO law on INGO operations in the five years since the passage of the law. We test whether and how INGO operational space under the ONGO law is influenced by organizations' issue areas. We argue that the less contentious the issue areas INGOs work in the more likely they are granted greater operational space, measured by the number of provinces they are legally authorized to work in. China provides an excellent setting for examining the phenomenon of closing space civic space, since the 2017 ONGO law hypothetically regulates both claim-making and service-providing NGOs, while the level of ambiguity of the law gives ample room for arbitrary discretion and conflicting attitudes by authorities. We test our argument by analyzing administrative data from 633 registered representative offices of INGOs registered in China between 2017–2021 and test our hypothesis with a multilevel Bayesian model. Our results speak to the broader literature on closing civic space and provide an empirical illustration of the practical effect of NGO restrictions on global civil society.
keywords:
  - international NGOs
  - civil society
  - authoritarianism
  - Chinese ONGO law
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/compassionate-clam>"
bibliography: references.bib
link-citations: true
---

@@

@@

TODO: New cool thing here is that we can look at de facto implementation in greater detail. Existing studies use things like V-Dem's core civil society index or civil society repression index to capture the general regulatory environment, but changes in those scores aren't necessarily tied to specific laws. Here we can see that differential enforcement/de facto/de jure stuff directly

In this paper, we first explore the ongoing phenomena of closing civic space and authoritarian regulation of civil society, examining how governments conceptualize the role of international civil society in their domestic politics and why governments select regulatory regimes for international NGOs. We then investigate how international NGOs have responded and adapted to more restrictive regulatory environments and theorize that a key component of INGO regulatory adaptation is *organizational flexibility*, or the ability to operate with greater latitude in host countries. We propose two key determinants of this flexibility: (1) organizational characteristics and (2) host government preferences and relationships TODO and their interplay/interaction. We illustrate this theory using the case of China TODO stats and case studies, mixed method, demonstrating how INGOs that address issues that align with government preferences enjoy the most organizational flexibility in the country.

@@

TODO: Meng - add brief explanations of the TA/RO pathways figure and articulate the "cooperative requirements part", i.e., seeking a psu or Chinese partner

TODO: Andrew - image of TA/RO pathways @fig-ro-ta-pathways

@@

@@

(Meng added)

@@

@ChaudhryHeiss:2022 and the divergence in de jure and de facto regulations---the implementation doesn't have to match the written law, and often the written law can serve as a "weapon hanging on the wall that never fires" that serves as a warning to potentially uncooperative NGOs [@Kozenko:2015]. See also [@ToeplerZimmerFrohlich:2020; @Plantan:2022]
INGO regulations are implemented and enforced differentially - difference in de jure and de facto. This differential enforcement is one mechanism to allow host governments to harness the aspects of INGO programming that align with their domestic policy preferences while limiting INGO behavior that could threaten regime stability.

@@

::: content-hidden
"We discuss the role of civil society organizations (CSOs) as agents of democratization and note the emergence of dual, at times apparently conflicting policy postures within authoritarian regimes (restriction and repression for some CSOs vs. financial support and opportunities for collaboration for others)." [@ToeplerZimmerFrohlich:2020]

Also, [@Plantan:2022] finds "that I find" evidence of selective implementation that reveals which groups are seen as threatening or beneficial."
:::

# Empirical expectations

@@

# Data and empirical approach

Main structure for these expectations + our analysis:

@@

@@

TODO: Meng - TA vs. RO, exploring main mechanisms of contentiousness, and relationships

@@

>>>>

---
title: "The Implementation of China’s Overseas NGO Law and the Operating Space for International Civil Society"
short-title: "China's ONGO Law and International Civil Society"
date: "2023-11-27"
author:
- name: Meng Ye
  affiliation: Georgia State University
  email: mye2@student.gsu.edu
  orcid: 0000-0002-6552-8338
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
abstract: |
  TODO: **Update this based on the finished version of the paper.** China’s 2017 Overseas NGO (ONGO) Law is part of a larger global trend of legal restrictions on international nongovernmental organizations (INGOs). However, the effects of this global crackdown on INGOs have been difficult to measure since there is often divergence between the formal de jure regulations and the de facto implementation of these laws. Additionally, the severity and consistency of the enforcement of these laws often depends on the type of services provided by these organizations. In this paper, we explore the effect of China’s ONGO law on INGO operations in the five years since the passage of the law. We test whether and how INGO operational space under the ONGO law is influenced by organizations' issue areas. We argue that the less contentious the issue areas INGOs work in the more likely they are granted greater operational space, measured by the number of provinces they are legally authorized to work in. China provides an excellent setting for examining the phenomenon of closing space civic space, since the 2017 ONGO law hypothetically regulates both claim-making and service-providing NGOs, while the level of ambiguity of the law gives ample room for arbitrary discretion and conflicting attitudes by authorities. We test our argument by analyzing administrative data from 633 registered representative offices of INGOs registered in China between 2017–2021 and test our hypothesis with a multilevel Bayesian model. Our results speak to the broader literature on closing civic space and provide an empirical illustration of the practical effect of NGO restrictions on global civil society.
keywords:
  - international NGOs
  - civil society
  - authoritarianism
  - Chinese ONGO law
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/compassionate-clam>"
bibliography: references.bib
link-citations: true
---

@@

TODO: Andrew - make this a paragraph, not a list.

@@

TODO: Andrew: Also, new cool thing here is that we can look at de facto implementation in greater detail. Existing studies use things like V-Dem's core civil society index or civil society repression index to capture the general regulatory environment, but changes in those scores aren't necessarily tied to specific laws. Here we can see that differential enforcement/de facto/de jure stuff directly

TODO: Andrew: Roadmap of paper, to add later
In this paper, we first explore the ongoing phenomena of closing civic space and authoritarian regulation of civil society, examining how governments conceptualize the role of international civil society in their domestic politics and why governments select regulatory regimes for international NGOs. We then investigate how international NGOs have responded and adapted to more restrictive regulatory environments and theorize that a key component of INGO regulatory adaptation is *organizational flexibility*, or the ability to operate with greater latitude in host countries. We propose two key determinants of this flexibility: (1) organizational characteristics and (2) host government preferences and relationships TODO and their interplay/interaction. We illustrate this theory using the case of China TODO stats and case studies, mixed method, demonstrating how INGOs that address issues that align with government preferences enjoy the most organizational flexibility in the country.

@@

TODO: Meng - add brief explanations of the TA/RO pathways figure in @fig-ro-ta-pathways and articulate the "cooperative requirements part", i.e., seeking a psu or Chinese partner

@@

TODO: Andrew: I'll clean up/rewrite this section and the section headings, etc.

@@

@@

TODO: Andrew - I'll clean up this section

INGO regulations are implemented and enforced differentially - difference in de jure and de facto. This differential enforcement is one mechanism to allow host governments to harness the aspects of INGO programming that align with their domestic policy preferences while limiting INGO behavior that could threaten regime stability.

@@

# Empirical expectations and analytical approach

TODO: Andrew: I'll clean up this section

@@

@@

TODO: Meng

@@

TODO: Meng - TA vs. RO, exploring main mechanisms of contentiousness and relationships

@@

::: {#refs .bibliography}
:::
